\LARGE \bf Mean Estimation from One-Bit Measurements

### 2022-04-20 introduction.tex
addressed specific commets in R2

In setting (i), the estimator can evaluate any optimal estimator of location
(e.g., the sample mean if the data is Gaussian), then quantize it using
$n$ bits. As the accuracy in describing the empirical mean decreases
exponentially in $n$, the quantization error is negligible compared
to the statistical error in mean estimation~\cite{720540}. That is,
centralized encoding induces no asymptotic efficiency loss.
%
The story is different in settings (ii) and (iii). Precisely, we show that
in the adaptive setting~(ii), the optimal efficiency of a one-bit scheme is
(asypmtotically) precisely that of the sample median, and that this
efficiency is achievable. As a concrete example, when $X_i$ are
i.i.d.\ Gaussian, we necessarily lose a factor of $\pi/2 \approx 1.57$ in
the asymptotic risk; the one-bit constraint decreases the effective sample
size by a factor of $\pi/2$ compared to estimating it without the bit
constraint. It turns out that, in the settings we consider,
only a \emph{single round} of adaptivity (see Fig.~\ref{fig:one_round} for
an illustration) is sufficient to achieve optimal convergence rates.
%
In distinction from setting (ii), in setting (iii) when the messages must be
independent, there is no distributed estimation scheme that achieves the
efficiency of the sample median uniformly over $\theta$.  We establish this
result via Le Cam's local asyptotic normality theory, allowing us
to provide exact characterizations of the asymptotic efficiency
of suitably regular encoding schemes.
%% To summarize our
%% contributions, we establish that the optimal ARE in settings (i) is $1$, the
%% optimal ARE in setting (ii) is ARE of the sample median, and no uniformly
%% optimal procedure exists in setting (iii).

>>>>

In setting (i), the estimator can evaluate any optimal estimator of location
(e.g., the sample mean if the data is Gaussian), then quantize it using
$n$ bits. As the accuracy in describing the empirical mean decreases
exponentially in the number of bits, the quantization error is negligible compared
to the statistical error in mean estimation~\cite{720540, cai2020distributed}. That is,
centralized encoding induces no asymptotic efficiency loss.
%
The story is different in settings (ii) and (iii). Precisely, we show that
in the adaptive setting~(ii), the optimal efficiency of a one-bit scheme is
(asypmtotically) precisely that of the sample median, and that this
efficiency is achievable. As a concrete example, when $X_i$ are
i.i.d.\ Gaussian, we necessarily lose a factor of $\pi/2 \approx 1.57$ in
the asymptotic risk; the one-bit constraint decreases the effective sample
size by a factor of $\pi/2$ compared to estimating it without the bit
constraint. It turns out that, in the settings we consider,
only a \emph{single round} of adaptivity (see Fig.~\ref{fig:one_round} for
an illustration) is sufficient to achieve optimal convergence rates.
%
In distinction from setting (ii), in setting (iii) when the messages must be
independent, there is no distributed estimation scheme that achieves the
efficiency of the sample median uniformly over $\theta$.  We establish this
result via Le Cam's local asyptotic normality theory, allowing us
to provide exact characterizations of the asymptotic efficiency
of suitably regular encoding schemes.
%% To summarize our
%% contributions, we establish that the optimal ARE in settings (i) is $1$, the
%% optimal ARE in setting (ii) is ARE of the sample median, and no uniformly
%% optimal procedure exists in setting (iii).

### 2022-04-20 onebit_full.tex
addressed specific commets in R2

\author{
    \IEEEauthorblockN{
Alon Kipnis and   
    John C. Duchi }

\thanks{
Alon Kipnis is with the School of Computer Science at Reichman University, Herzliya, Israel (kipnisal@stanford.edu).} 
\thanks{J. Duchi is with the Department of Statistics and Department of Electrical Engineering at Stanford University, Stanford, CA, 94035 (jduchi@stanford.edu).}
\thanks{Copyright (c) 2017 IEEE. Personal use of this material is permitted.  However, permission to use this material for any other purposes must be obtained from the IEEE by sending a request to pubs-permissions@ieee.org.}
\thanks{This paper was presented in part at the 55th Annual Allerton Conference on Communication, Control, and Computing (Allerton) \cite{KipnisAllerton2017}. }
\thanks{
The work of A. Kipnis was supported in part by funding from the NSF under Grant No.~DMS-1418362 and DMS-1407813, and by a fellowship from the Koret Foundation.}
}

@@

>>>>

\author{
    \IEEEauthorblockN{
Alon~Kipnis and   
    John~C.~Duchi}

\thanks{
Alon Kipnis is with the School of Computer Science at Reichman University, Herzliya, Israel (alon.kipnis@idc.ac.il).} 
\thanks{J. Duchi is with the Department of Statistics and the Department of Electrical Engineering at Stanford University, Stanford, CA, 94035 (jduchi@stanford.edu).}
\thanks{Copyright (c) 2017 IEEE. Personal use of this material is permitted.  However, permission to use this material for any other purposes must be obtained from the IEEE by sending a request to pubs-permissions@ieee.org.}
\thanks{This paper was presented in part at the 55th Annual Allerton Conference on Communication, Control, and Computing (Allerton) \cite{KipnisAllerton2017}. }
\thanks{
The work of A. Kipnis was supported in part by funding from the NSF under Grant No.~DMS-1418362 and DMS-1407813, and by a fellowship from the Koret Foundation.}
}

@@

\begin{IEEEbiographynophoto}{Alon Kipnis}
\input{ak-bio}
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{Alon Kipnis}
\input{jd-bio}
\end{IEEEbiographynophoto}
